---
title: "Linear mixed-effects models"
description: |
  Effect of feeding frequency on juvenile coral growth
date: "2025-04-14"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: Show code
---

# Overview

This code analyzes measurement data of juvenile *Orbicella faveolata* and *Pseudodiploria clivosa* colonies in an ex-situ nursery and investigates the effect of amino bath feeding frequency on coral growth.

Research conducted at The Reef Institute in West Palm Beach, FL. Thanks to Jasmine Rios for data collection.

# Setup

First, we will set document-wide options and imports all necessary packages.

```{r setup, results = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# import packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(lme4)
library(lmerTest)
library(emmeans)
library(lubridate)
library(modelsummary)
```

# Data input

Next, we will import the raw growth measurements data.

```{r}
measurements <- read_xlsx(here("data", "Feeding study measurements.xlsx"))
```

# Data wrangling

Then we can format the data to be more useful for analysis.

```{r}
# format date column and add columns for treatment group and species
measurements_clean <- measurements %>% 
  filter(!is.na(`Area (mm^2)`)) %>% 
  mutate(`Date of picture` = date(`Date of picture`)) %>% 
  mutate(treatment_group = substr(`Combined group name`, 1, 1)) %>% 
  mutate(species = case_when(endsWith(`Combined group name`, "O") ~ "OFAV",
                             endsWith(`Combined group name`, "P") ~ "PCLI")) %>% 
  mutate(combined_individual_name = paste0(`Combined group name`, "-", Individual)) %>% 
  mutate(days_elapsed = as.numeric(`Date of picture` - date("2025-03-04")))

# cast treatment group as a factor, set group 3 (control) as reference
measurements_clean$treatment_group <- factor(measurements_clean$treatment_group, levels = c("3", "2", "1"))

# create data frame with only PCLI measurements
pcli_measurements <- measurements_clean %>% 
  filter(species == "PCLI")

# create data frame with only OFAV measurements
ofav_measurements <- measurements_clean %>% 
  filter(species == "OFAV")

# calculate average sizes for each treatment group, measurement date, and species
avg_sizes_by_group <- measurements_clean %>% 
  group_by(`Combined group name`, `Date of picture`, treatment_group, species) %>% 
  summarize(mean_area = mean(`Area (mm^2)`), se = sd(`Area (mm^2)`) / sqrt(n()))
```

# Data visualization

This code creates line graphs showing the individual measurements and average measurements for each treatment group over time.

```{r}
ggplot(measurements_clean, aes(x = `Date of picture`, y = `Area (mm^2)`, color = combined_individual_name)) +
  geom_line() +
  facet_wrap(~ species, scales = "free") +
  scale_y_continuous(limits = c(0, NA)) +
  theme(legend.position = "none")
```

Figure 1. Growth of juvenile *Orbicella faveolata* (OFAV) and *Pseudodiploria clivosa* (PCLI) coral colonies over the duration of the study.

```{r}
ggplot(avg_sizes_by_group, aes(x = `Date of picture`, y = mean_area, color = treatment_group, fill = treatment_group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = mean_area - se, ymax = mean_area + se), alpha = 0.2, color = NA) +
  facet_wrap(~ species, scales = "free") +
  scale_y_continuous(limits = c(0, NA)) +
  labs(y = "Mean area (mm^2)", color = "Treatment group") +
  guides(fill = "none")
```

Figure 2. Effects of amino bath feeding on the growth of juvenile *Orbicella faveolata* (OFAV) and *Pseudodiploria clivosa* (PCLI) coral colonies, divided by treatment group, over the duration of the study. Group 1 was fed weekly, group 2 was fed every two weeks, and group 3 was not fed at all. Lines represent the mean size of corals, while shaded regions represent the mean ± standard error. Feeding had little to no effect on OFAV colonies, while more frequent feeding resulted in faster growth for PCLI colonies.

# Linear mixed-effects models

Finally, this code creates linear mixed-effects models to investigate the effect of feeding frequency on coral growth quantitatively.

Table 1. Summary of linear mixed-effects models of the effect of feeding frequency on growth of juvenile *Orbicella faveolata* (OFAV) and *Pseudodiploria clivosa* (PCLI) coral colonies. Initial colony sizes were treated as random effects, while the number of days elapsed and treatment group were treated as fixed effects. Feeding frequency had no significant effect on the growth rate of OFAV colonies (*p* > 0.1), while it had a positive effect on the growth rate of PCLI colonies, falling just below the threshold of significance (0.05 < *p* < 0.1).

```{r}
pcli_model <- lmer(`Area (mm^2)` ~ days_elapsed * treatment_group + (1 | combined_individual_name), data = pcli_measurements)

ofav_model <- lmer(`Area (mm^2)` ~ days_elapsed * treatment_group + (1 | combined_individual_name), data = ofav_measurements)

modelsummary(list("OFAV" = ofav_model, "PCLI" = pcli_model),
             output = "html",
             stars = TRUE)
```
